---
title: "Report of PSID1982 Data Set"
author:"Burak Bagatur"
output: html_document
date: "2024-12-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This report includes the analyzes and results performed on the data set.


 Data Set:PSID1982 
WAGE: Wage  paid for an hour (USD), Class: Float
EXPERIENCE: Work experience Class: Intiger
weeks:Weeks worked in a year. 
Occupation: Departments in the business department , Class:Character
Ethnictiry : Race/ethnicity of the individual
Industry: Industry type
sOUTH: Region: Southern USA
SMSA: Standard Metropolitan Statistical Area (Urban vs. Rural)
Union: Union membership , Class: Factor
Gender: whether male or female, Class:Character
Married: Marital status, Class: Factor
EDUCATION: Years of education completed.



```{r PSID1982}
df<- PSID1982
str(df)
summary(df )
```

## Including Plots

Histogram of Wage and experience
boxplot of wage~gender 

```{r pressure, echo=FALSE}
hist(df$wage , main ="Wage Histogram ", xlab="Wage",col="yellow")
hist(df$experience, main ="Experience Histogram",xlab="Experience",col="brown")
boxplot(wage ~ gender, data=df,main="Boxplot of wage and gender",xlab="gender",ylab="wage ",col=c("white","grey"))

```

#Regresion Analysis 

```{r setup, include=FALSE}
##Regresion Analysis result 
###The factors affecting wages are examined with the regression model below.
reg_model1 <- lm(wage~experience+education+gender+union, data=df )
summary(reg_model1)
###The factors affecting education  is  examined with the regression model below.
glm_model1 <- glm(education ~experience, data=df,family=gaussian("identity"))
summary(glm_model1)

df$union <- ifelse(df$union =="yes",1,0)
glm_model3 <- glm(union ~wage ,data=df, family=binomial(link="logit"))
summary(glm_model3)

glm_model_poisson <- glm(experience ~  married + education+wage  , data = df, family = poisson(link="log"))
summary(glm_model_poisson)
```